### 2016-01-12 proposal/proposal.tex
initial commit with proposal directory and placeholder tex file for proposal

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Nass Replication Study}
\author{Matthew Neal, Shaown Sarker}
\maketitle
\end{document}