### 2016-03-21 paper/draft1.tex
Renamed "sig..." to draft1.tex.  Set up CI.

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012
\documentclass{sig-alternate-05-2015}

\usepackage{comment}    % For block comments
\usepackage{cite}       % For group citations
\usepackage{color}      % For warnings and draft notations

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{}

% ISBN
\isbn{}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}
%\acmPrice{\$15,000}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2016} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Software Developers and Computers as Social Actors}
%\subtitle{A Nass Replication}
%

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{Prairie Rose Goodwin \qquad Adam Marrs \qquad Matthew Neal \qquad Shaown Sarker\\ \\ \affaddr{Department of Computer Science}\\ \affaddr{North Carolina State University}\\ \email{\normalsize \{prgoodwi,acmarrs,mneal,ssarker\}@ncsu.edu}\\}

\maketitle
\begin{abstract}
\begin{itemize}
    \item Standard abstract, defines what we were able to achieve.
\end{itemize}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003121</concept_id>
<concept_desc>Human-centered computing~Human computer interaction (HCI)</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Human-centered computing~Human computer interaction (HCI)}

%
%  Use this command to print the description
%
\printccsdesc

\keywords{software engineering; computers as social actors}

\section{Introduction}
\begin{itemize}
    \item Define what Nass' study was.
    \item Define what is novel about our study in comparison to Nass' original study.
    \item Introduce who our subjects are and the basics of what our methodology was.
\end{itemize}

\section{Related Work}

\begin{comment}
In research interviews open, frank speech is essential to the process. Interviews are plagued by biases inherent in social interactions. Such biases include the normative-response bias (people not wanting to admit to answers that are counter to the perceived norm), interviewer-based bias (people answering based on the perceived preference of the interviewer), and politeness norms (people modifying answers to avoid offending). Computer systems were introduced to mitigate these known social-biases by providing anonymity. However, through a series of studies, Nass et al. documented how individuals anthropomorphize technology and exhibit attitudes previously thought to be reserved for person to person interaction \cite{nass1999people}\cite{reeves1996people}. Contrary to what was believed at the time, the results suggested that humans unintentionally, unknowingly apply social rules to computer systems.   
This foundational work sparked further studies to replicate, validate, and extend this assertion. Since the initial set of studies were conducted entirely in the United States, additional research validated the findings across other cultures \cite{katagiri2001cross}. Furthermore, the existence of computers as social entities has been studied to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}. To our knowledge, no studies currently exist which attempt to replicate these effects with software developers.   
In order to create better software, it is important to study the attitudes of software developers and how they view their work environments. Given software developers’ advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present. At the same time, it has been shown that humans with accurate mental models are more effective in maintaining teammate bonds with automated computer systems \cite{wilkison2008effects}. Given this premise, we are interested in exploring the following questions:

\begin{itemize}
    \item Do software developers exhibit social-biases towards computers?
    \item Does ownership1 of the computer modify the exhibited social-biases? 
\end{itemize}
\end{comment}

Previous research demonstrates that computers are often treated as social entities despite their inanimate nature. There have been studies that focus on the human application of social norms and expectations to computers in a seamless manner. Nass et al. performed a series of experiments that pointed out that social stereotypes attributed to gender, ethnicity, and group loyalty, along with deeply ingrained social habits and behaviors like, reciprocity and reciprocal self-disclosure, are extended to computers when interacting with them~\cite{Nass2000machines,nass1999people}. Moreover, prior studies documented that this behavior is not only limited to computers, human users anthropomorphize other technological mediums like television~\cite{reeves1996people}.

\hfill

These findings engendered a number of studies that extended on this implication by replicating, validating, and exploring the possible factors behind these kind of responses. The existing Computers as Social Actors (CASA) studies featured an implicit bias - although the studies were conducted with participants exclusively from the United States, the findings were generalized to all cultures. Katagiri and colleagues performed a comprehensive study to determine whether social rules derived from different cultures affect human users responses to computers by observing participants from the United States and Japan~\cite{katagiri2001cross}. Furthermore, researchers studied the existence of computers as social entities to better understand racial biases in interviews\cite{krysan2003race}, how the quality of interaction affects expressive systems \cite{vidyarthi2011sympathetic}, and even how to leverage social cues to more effectively sell products online \cite{wang2007can}.

\hfill

Software developers maintain a collaborative environment with their development machines to achieve their goals. Given software developers' advanced understanding of how computers work, one would expect the attribution of social qualities to computers to be minimal or not present.  This type of attribution would be expected, if the the sociability of a computer system is modified by the quality of an individual's mental model of a computer system. Research has shown that human-computer team performance is higher when human users have accurate mental models~\cite{wilkison2008effects}. Further studies by Bickmore and Picard have documented the long-term relation establishment and maintenance between a human-computer team by observing that human users of a relational computer agent did not only respect, trust, and like the agent, but also wanted to continue working with the agent even after the end of the study~\cite{Bickmore:2005:EML:1067860.1067867}.

\hfill

In this paper, we investigate whether a long term relationship between a software developer and her computer has any significant effect on the social norm of politeness witnessed by Nass et al in their original study~\cite{nass1999people} by contrasting their responses to computers with that of non-software developers.

Our work explores the effect of viewing computers as a human-like social entity on the very specific demographic of software developers and attempts to determine the level of politeness software developers exhibit toward their machines.

\section{Methods}

\subsection{Participants}
Here we will detail the composition of our study participants including:
\begin{itemize}
    \item The number in each group
    \item Programming experience breakdown
    \item Gender breakdown (potentially)
\end{itemize}

\subsection{Procedure}
Generally, here we will detail the procedure by which the study was conducted.

\begin{itemize}
    \item Screenshots of the website which participants visited.
    \item Detailed information about the experimental environment: 
    \begin{itemize}
        \item Where the study took place.
        \item What machines people used.
        \item How long the study took.
        \item Explaination of the function of the website used.
    \end{itemize}
\end{itemize}

\section{Results}

\begin{itemize}
    \item Overall description of what our results were.
\end{itemize}

\subsection{Analysis}

\begin{itemize}
    \item Graph describing the differences between CSC group and Psychology group, along with discussion of that graph.
    \item Table describing differences between CSC groups and Psych group, with discussion.
    \item Appropriate statistical tests to describe potential inferences that can be made from the results.
\end{itemize}

\section{Discussion}

\begin{itemize}
    \item Overall discussion of the importance or lack of importance of our analysis and results.
\end{itemize}

\subsection{Threats to Validity}

\subsubsection{Internal Validity}

\begin{itemize}
    \item Describe the mistakes we made.
\end{itemize}

\subsubsection{External Validity}

\begin{itemize}
    \item Describe whether our results are significant enough to generalize to future studies.
\end{itemize}

\subsection{Future Work}

\begin{itemize}
    \item Describe any potential studies that we believe can build on our own.
\end{itemize}

\section{Conclusion}
Although I'm not a super big fan of conclusions, it makes sense to have something short here I'd think.

\section*{Acknowledgements}
We would like to thank...

\bibliographystyle{abbrv}
\bibliography{sigproc}

\appendix
\section{Website and Survey Transcripts}

\end{document}